---
title: "ToothGrowth Data Analysis"
author: "Taedong Yun"
output: html_document
---

## Overview

In this article, we will analyze the ToothGrowth data in the R datasets package. We will perform basic exploratory data analyses by providing a summary of the data and elementary statistical inference.

## Summary of Data

The ToothGrowth dataset contains the result of an experiment on Guinea Pigs about the correlation between Vitamin C dosage and the growth of tooth. Let us take a first look of the dataset.

```{r}
library(datasets)
str(ToothGrowth)
```

The dataset has three variables, `len`, `supp`, and `dose`. `len` is the response which is the length of a tooth, `supp` is a delivery type of Vitamin C, swhich is either "VC" (ascorbic acid) or "OJ" (orange juice), and `dose` is the amount of dosage in milligrams. Note that `dose` variable take a value in 0.5, 1, or 2 so we can make it a factor variable. This dataset contains 60 observations. More information about this dataset can be found at https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/ToothGrowth.html. Here is a basic summary of each variables in the dataset.

```{r}
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
summary(ToothGrowth)
```

Note that `supp` and `dose` factor variables are evenly distributed in their possible values. The following figure is a box plot of the dataset.

```{r echo=FALSE, fig.height = 3}
library(lattice)
bwplot(len ~ supp | dose, data = ToothGrowth, ylab = "Length", xlab = "Types of Supplement & Dosage", main = "Data Summary")
```

## Inference

From the boxplot in the previous section, we study the following the problems.

1. Do two different types of supplement have different effects on the growth of a tooth?
2. Does more dosage means more growth of a tooth?

We will address these two questions with statistical inference techniques.

### Length vs Supplement Type



### Length vs Dosage



## Conclusion

*INCLUDE ASSUMPTIONS NEEDED*
